---
title: "Award Voting Behavior"
author: "Alex Novet"
date: "April 28, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(rvest)
require(janitor)
require(ggthemes)
require(tidyverse)
```

To start, scrape the main awards data from each year's page on Hockey Reference

```{r scrape hockeydb}

#year <- 2017
#url <- read_html(paste0("https://www.hockey-reference.com/awards/voting-", year, ".html"))

  

years <- c(2007:2017)
awards <- c("selke", "norris", "vezina", "byng")
dat1 <- data.frame()
for (j in awards){
  for (i in years){
    url <- read_html(paste0("https://www.hockey-reference.com/awards/voting-", i, ".html"))
  
    temp <- url %>% 
      html_nodes(xpath = '//comment()') %>%    # select comment nodes
      html_text() %>%    # extract comment text
      paste(collapse = '') %>%    # collapse to a single string
      read_html() %>%    # reparse to HTML
      html_node(paste0('table#',j,'_stats')) %>%    # select the desired table
      html_table() %>%
      mutate(year = i, award = j)
    dat1 <- bind_rows(dat1, temp)
  }
}

```

The calder table has 2 header rows, which throws things off. This could be done more efficient, but it's easiest to redo the pull with just the Calder and then clean it up

```{r calder}
calder <- data.frame()
for (i in years){
  url <- read_html(paste0("https://www.hockey-reference.com/awards/voting-", i, ".html"))

  temp <- url %>% 
    html_nodes(xpath = '//comment()') %>%    # select comment nodes
    html_text() %>%    # extract comment text
    paste(collapse = '') %>%    # collapse to a single string
    read_html() %>%    # reparse to HTML
    html_node("table#calder_stats") %>%    # select the desired table
    html_table()
  colnames(temp) <- temp[1,]
  temp <- tail(temp,-1) %>% mutate(year = i, award = "calder")
  calder <- bind_rows(calder, temp)
}

char <- select(calder, Player, Tm, Pos, award)
num <- select(calder, -Player, -Tm, -Pos, -award) %>% mutate_all(as.numeric)
calder2 <- cbind(char, num)

```

Also need a separate pull for Hart becuase that table is saved differently on the web page


```{r hart}
hart <- data.frame()
for (i in years){
  url <- paste0("https://www.hockey-reference.com/awards/voting-", i, ".html")
  temp <- url %>% read_html() %>% html_table(fill = TRUE)
  temp <- temp[[1]]
  colnames(temp) <- temp[1,] 
  temp <- tail(temp, -1) %>% mutate(year = i, award = "hart")
  hart <- bind_rows(hart, temp)
}

char <- select(hart, Player, Tm, Pos, award)
num <- select(hart, -Player, -Tm, -Pos, -award) %>% mutate_all(as.numeric)
hart2 <- cbind(char, num)

```

Now we just need to combine the datasets and then we can start pulling stats. There's lots of different ways to measure voting behavior, most of which involve some pretty arbitrary cuts. I think ones I'm particularly interested in are:

1. Number of people receiving any votes for the award (regardless of whether they were first, second, etc.)
2. Players with 5 or fewer votes

```{r clean}


dat2 <- bind_rows(dat1, calder2) %>% bind_rows(hart2) %>% clean_names() %>%
  mutate(x4th = ifelse(is.na(x4th), 0, x4th),
         x5th = ifelse(is.na(x5th), 0, x5th)) #Replace NAs in Vezina voting
save(dat2, file = "Award Voting Annual Data.Rds")


stats <- group_by(dat2, award, year) %>%
  summarise(total_players = n(),
            few_votes = sum((x1st +x2nd + x3rd +x4th +x5th) <= 5))

new <- data.frame(award = c("hart", "norris", "vezina", "calder", "selke", "byng"),
                  year = 2018,
                  total_players = c(20,18,10,12,31,56),
                  few_votes = c(10,7,6,2,21,36))
stats <- bind_rows(stats, new)

```


```{r viz}

ggplot(stats, aes(x = year, y = total_players, group = award)) + 
  geom_line(aes(color = award), size = 1.5) + geom_vline(xintercept = 2017) +
  labs(title = "Players Receiving an Award Vote", x = "Year", y = "Number of Players") +
  theme_few()

ggplot(filter(stats, year != 2014), aes(x = year, y = total_players, group = award)) + 
  geom_line(aes(color = award), size = 1.5) + geom_vline(xintercept = 2017) +
  labs(title = "Players Receiving an Award Vote", x = "Year", y = "Number of Players") +
  theme_few()
  
ggplot(stats, aes(x = year, y = few_votes, group = award)) + 
  geom_line(aes(color = award), size = 1.5) + geom_vline(xintercept = 2017) +
  labs(title = "Players Receiving 5 or Fewer Award Votes", x = "Year", y = "Number of Players") +
  theme_few()

ggplot(filter(stats, year != 2014), aes(x = year, y = few_votes, group = award)) + 
  geom_line(aes(color = award), size = 1.5) + geom_vline(xintercept = 2017) +
  labs(title = "Players Receiving 5 or Fewer Award Votes", x = "Year", y = "Number of Players") +
  theme_few()
```

Thanks to hockey-reference for the data. Note that this doesn't take into consideration changes in the # of total voters